Journal and article coverage

### 2025-02-19 analysis/manuscript.qmd
Starting publisher analysis. Focus: overrep of big 3, newly launched springer hybrid oa only in hoaddata, lacking corssref license practise in some other publishers (name and shame)

@@

Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only available via hoaddata (grey area).
After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews in hoaddata. Particulalry, Letters and Editorials could not be fully detected.
Paratext recognition also failed in some cases to identify non-scientific content such as journal covers or reviewer lists, which are generally not covered by Scopus and Web of Science. 
To a lesser extent, differences in publication and indexing dates were a reason for non-overlapping DOIs.

Using DOIs also the size of journals were calculated (Fig C). It highlights thate large spread in the intersection between all three datasett. Journals are, on average, smaller than those just covered by one or two sources.In particular journals exlcusively covered by hoaddata are significant smaller than those covered by all three sources.

>>>>

### Overview

@@

```{r miss_dois_check}
miss_dois <- readr::read_csv(here::here("data-raw", "missing_dois_in_active_jns.csv"))
miss_dois_val <- readr::read_csv(here::here("data-raw", "miss_dois_in_active_jns_ckeck.csv"))
```
Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only available via hoaddata (grey area).
In case of hybrid ournals covered by all three data sources, `r miss_dois |> nrow()` DOIs were only present in hoaddata.
After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews in hoaddata. Particulalry, letters and editorials could not be fully detected.
Paratext recognition failed in `r round(miss_dois_val |> filter(is.na(scp_item_type), is.na(wos_item_type)) |> pull(n_doi) / miss_dois |> nrow() * 100)`% of DOIs to identify non-scientific content such as front matters or reviewer lists, which are generally not indexed by Scopus and Web of Science. 
To a lesser extent, differences in publication and indexing dates were a reason for non-overlapping DOIs.

Using overlapping DOIs, the publication volume between 2019 and 2023 was also calculated for each journal. 
Figure \ref{fig-upset_coverage_results}C shows the distribution for each combination. It shows that there is a large spread across the journals covered by all three data sources. 
Furthermore, journals in this set published more on average than journals covered by only one or two data sources. 
In particular, the journals covered exclusively by hoaddata are substantially smaller than those covered by all three sources. 
Upon inspection, these are often newly launched journals, which explains the relatively low five-year publication volume.

### Coverage by publisher portfolio

```{r publisher_coverage_data_prep}
# Add ESAC portfolio names
esac <- hoaddata::jct_hybrid_jns |>
  distinct(issn_l, esac_publisher) |>
  mutate(
    publisher = case_when(
      esac_publisher == "Elsevier" ~ "Elsevier",
      esac_publisher == "Springer Nature" ~ "Springer Nature",
      esac_publisher == "Wiley" ~ "Wiley",
      is.character(esac_publisher) ~ "Other"
    )) |>
  mutate(publisher =
           forcats::fct_relevel(
             publisher,
             c("Elsevier", "Springer Nature", "Wiley", "Other") 
           )) |>
  mutate(publisher = forcats::fct_rev(publisher)) |>
  distinct(issn_l, publisher)

jn_upset_publisher <- inner_join(upset_data, esac, by = "issn_l") |>
  # Because of Journal transfer, restrict choice to one portfolio
  distinct(issn_l, .keep_all = TRUE)
```

Figre shows the coverage of hybrid journals in transformative agreements by publisher portfolio according to the ESAC registry.

```{r upset_publisher_plot, fig.asp=0.75}
size = get_size_mode('exclusive_intersection')

upset_publisher_plot <- ComplexUpset::upset(
  jn_upset_publisher,
  c("Crossref", "WoS", "Scopus"),
  # Only journals in Crossref
  min_size = 30,
  width_ratio = 0.4,
  stripes = "transparent",
  name = "Data Source Coverage",
  base_annotations = NULL,
  annotations = list(
    "Journal size\n(log-scale)" = (
      ggplot(mapping = aes(x = intersection, y = overlap, fill = fct_rev(publisher))) +
        geom_boxplot(outliers = FALSE, width = bar_width) +
        scale_fill_manual(
          "",
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3"
          ), guide = "none") +
        scale_y_log10(labels = scales::number_format(big.mark = ",")) +
        ggtitle("C")
    ),
    "Articles" = (
      ggplot(mapping = aes(y = overlap, fill = publisher, color = publisher)) +
        geom_bar(stat = "identity", width = bar_width) +
        scale_fill_manual(
          "",
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3a0"
          ), guide = "none") +
        scale_color_manual(
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3"
          ), guide = "none"
        ) + 
        scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
        ggtitle("B") 
    ),
    "Publisher" = (
      ggplot(mapping = aes(fill = publisher)) +
        geom_bar(stat = "count", position = "fill", width = bar_width) + 
        scale_y_continuous("Journals", labels = scales::percent_format()) +
        scale_fill_manual(
          "",
          values = c(
            "Elsevier" = "#e9711c",
            "Springer Nature" = "#486a7e",
            "Wiley" = "#068853",
            "Other" = "#b3b3b3"
          ), guide = "none") +
        ggtitle("A")
    )
  ),
  set_sizes = (
    upset_set_size(
      geom = geom_bar(aes(fill = publisher), width = 0.8), 
      position = "right"
    ) +
    scale_fill_manual(
      "",
      values = c(
        "Elsevier" = "#e9711c",
        "Springer Nature" = "#486a7e",
        "Wiley" = "#068853",
        "Other" = "#b3b3b3"
      )
    ) +
    scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
    guides(fill = guide_legend(reverse = TRUE)) +
    ylab("Journals")
  ),
  guides = "over",
  matrix = (
    intersection_matrix(
      geom = geom_point(
        shape = "square",
        size = 2
      )
    )
  )
)

upset_publisher_plot & my_upset_theme 
```

### 2025-02-19 analysis/manuscript.tex
Starting publisher analysis. Focus: overrep of big 3, newly launched springer hybrid oa only in hoaddata, lacking corssref license practise in some other publishers (name and shame)

@@

In terms of article coverage, Figure \ref{fig-upset_coverage_results}B
shows the total publication volume per combination in terms of DOI
availability. The largest set of hybrid journals, which includes all
three data sources, also contains the largest number of articles. In
total, these journals recorded 6,289,687 overlapping articles,
represented by the blue bar. They represent 94\% of core articles with
DOI indexed in the Web of Science sample, and 86\% in Scopus. Another
657,697 articles are exclusive to both Scopus and hoaddata. Exclusively
in hoaddata were 177,110 articles, and exclusively in the proprietary
databases were 325,194 articles.

Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only
available via hoaddata (grey area). After validation at the DOI level
using the KB databases and manual inspection, the main reasons for
missing DOI coverage in the proprietary database were insufficient
classification of journal content as original research articles and
reviews in hoaddata. Paratext recognition also failed in some cases to
identify non-scientific content such as journal covers or reviewer
lists, which are generally not covered by Scopus and Web of Science. To
a lesser extent, differences in publication and indexing dates were a
reason for non-overlapping DOIs.

Using DOIs also the size of journals were calculated (Fig C). It
highlights thate large spread in the intersection between all three
datasett. Journals are, on average, smaller than those just covered by
one or two sources.In particular journals exlcusively covered by
hoaddata are significant smaller than those covered by all three
sources.

>>>>

\subsubsection{Overview}\label{overview}

@@

In terms of article coverage, Figure \ref{fig-upset_coverage_results}B
shows the total publication volume per combination in terms of DOI
availability. The largest set of hybrid journals, which includes all
three data sources, also contains the largest number of articles. In
total, these journals recorded 6,289,687 overlapping articles,
represented by the blue bar. They represent 94\% of core articles with
DOI indexed in the Web of Science sample, and 86\% in Scopus. Another
657,697 articles were exclusive to both Scopus and hoaddata. Exclusively
in hoaddata were 177,110 articles, and exclusively in the proprietary
databases were 325,194 articles.

Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only
available via hoaddata (grey area). In case of hybrid ournals covered by
all three data sources, 1,023,882 DOIs were only present in hoaddata.
After validation at the DOI level using the KB databases and manual
inspection, the main reasons for missing DOI coverage in the proprietary
database were insufficient classification of journal content as original
research articles and reviews in hoaddata. Particulalry, letters and
editorials could not be fully detected. Paratext recognition failed in
37\% of DOIs to identify non-scientific content such as front matters or
reviewer lists, which are generally not indexed by Scopus and Web of
Science. To a lesser extent, differences in publication and indexing
dates were a reason for non-overlapping DOIs.

Using overlapping DOIs, the publication volume between 2019 and 2023 was
also calculated for each journal. Figure
\ref{fig-upset_coverage_results}C shows the distribution for each
combination. It shows that there is a large spread across the journals
covered by all three data sources. Furthermore, journals in this set
published more on average than journals covered by only one or two data
sources. In particular, the journals covered exclusively by hoaddata are
substantially smaller than those covered by all three sources. Upon
inspection, these are often newly launched journals, which explains the
relatively low five-year publication volume.

\subsubsection{Coverage by publisher
portfolio}\label{coverage-by-publisher-portfolio}

Figre shows the coverage of hybrid journals in transformative agreements
by publisher portfolio according to the ESAC registry.

\begin{center}
\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/upset_publisher_plot-1.png}
\end{center}
